\begin{thebibliography}{99}

\bibitem{thecell} 
Alberts B et al. Molecular Biology of the Cell (4th edition)
Garland Pub (2002/03)

\bibitem{essentialtext}
Alberts B et al., 中村桂子、松原謙一監訳
「Essential細胞生物学(原著第2版)」 南江堂(2005/09)

\bibitem{iden} 
G. エドリン, Gordon Edlin(著), 伊藤 文昭, 大竹 英樹, 
蓑島 伸生, 井口 義夫, 清水 淑子, 清水 信義(翻訳)
「ヒトの遺伝学」東京化学同人 (1992/04)

\bibitem{ncRNA1}
河合剛太、金井昭夫(編)
「機能性Non-coding RNA」
クバプロ(2006/04)

\bibitem{pyrro}
Zhang Y, Baranov PV, Atkins JF, Gladyshev VN (2005)
Pyrrolysine and selenocysteine use dissimilar decoding strategies.
J Biol Chem. 280(21):20740-51.


\bibitem{myco} 
Fraser CM et al.(1995) 
The minimal gene complement of {\it Mycoplasma genitalium}.
Science 270(5235):397-403.

\bibitem{hs1}
Lander ES et al.(2001)
Initial sequencing and analysis of the human genome.
Nature. 409(6822):860-921. 

\bibitem{hs2}
Venter JC et al.(2001)
The sequence of the human genome.
Science. 291(5507):1304-51.

\bibitem{BI}
冨田勝(監修)
斎藤輪太郎(著)
「バイオインフォマティクスの基礎
〜 ゲノム解析プログラミングを中心に 〜」
サイエンス社(2005/07)

\bibitem{BLAST} Altschul SF et al.(1990) Basic local alignment search tool. 
J Mol Biol 215(3):403-10

\bibitem{BLAT} Kent WJ.(2002)
BLAT--the BLAST-like alignment tool.
Genome Res. 12(4):656-64. 

\bibitem{glang}
Arakawa K, Mori K, Ikeda K, Matsuzaki T, 
Kobayashi Y, Tomita M(2003)
G-language Genome Analysis Environment: a workbench for nucleotide 
sequence data mining.
Bioinformatics 19(2):305-306 http://www.g-language.org/

\bibitem{metab}
冨田勝、西岡孝明(編)
「メタボローム研究の最前線」
シュプリンガー・フェアラーク東京
(2003/11)

\bibitem{ac2}
津村ゆかり著
「よくわかる最新分析化学の基本と仕組み」
秀和システム(2009/4)

\bibitem{mass}
志田保夫、笠間健嗣、黒野定、高山光男、高橋利枝
「これならわかるマススペクトロメトリー」
化学同人(2001/03)

\bibitem{ac}  日本分析機器工業会編
「よくわかる分析化学のすべて」
日刊工業新聞社(2001/10)

\bibitem{iabsysbio}
Ishii N, Nakahigashi K, Baba T, Robert M, Soga T, Kanai A, Hirasawa T, 
Naba M, Hirai K, Hoque A, Ho PY, Kakazu Y, Sugawara K, Igarashi S, 
Harada S, Masuda T, Sugiyama N, Togashi T, Hasegawa M, Takai Y, Yugi K, 
Arakawa K, Iwata N, Toya Y, Nakayama Y, Nishioka T, Shimizu K, Mori H, 
Tomita M(2007)
Multiple High-Throughput Analyses Monitor the Response of {\it E. coli} to Perturbations
Science 316(5824):593-7


\bibitem{microar}
DeRisi JL, Iyer VR, Brown PO.(1997)
Exploring the metabolic and genetic control of gene expression on a genomic scale.
Science 278(5338):680-6. 

\bibitem{PPI}
Uetz P et al.(2000)
A comprehensive analysis of protein-protein interactions in {\it Saccharomyces cerevisiae}.
Nature 403(6770):623-7. 

\bibitem{thr}
Ra\"is B, Chassagnole C, Mazat JP.(1995)
Control of threonine pathway in {\it E. coli}. Application to biotechnologies.
Acta Biotheor. 43(4):285-97.

\bibitem{Ecell}
Tomita M, Hashimoto K, Takahashi K, Shimizu TS, Matsuzaki Y, Miyoshi F, Saito K, Tanida S, Yugi K, Venter JC, Hutchison CA 3rd.(1999)
E-CELL: software environment for whole-cell simulation.
Bioinformatics 15(1):72-84. http://www.e-cell.org/

\bibitem{Ecell3}
Takahashi K, Kaizu K, Hu B, Tomita M(2004) 
A multi-algorithm, multi-timescale 
method for cell simulation.
Bioinformatics. 20(4):538-46

\bibitem{Ginfo}
高木利久、冨田勝(編集)
「ゲノム情報生物学」
中山書店(2000/10) 

\bibitem{blood} Nakayama Y, Kinoshita A, Tomita M.(2005)
Dynamic simulation of red blood cell metabolism and its application to the analysis of a pathological condition.
Theor Biol Med Model 2(1):18. 

\bibitem{eXpanda} Negishi Y., Nakamura H., Yachie N., Saito R., Tomita M.
(2007) eXpanda: an integrated platform for network analysis and visualization. In Silico Biology 7, 0013. 
http://medcd.iab.keio.ac.jp/expanda/

\bibitem{cytoscape1} 
Shannon P, Markiel A, Ozier O, Baliga NS, Wang JT, Ramage D, Amin N, Schwikowski B, Ideker T.(2003)
Cytoscape: a software environment for integrated models of biomolecular interaction networks.
Genome Res. 2003 13(11):2498-504.

\bibitem{ecell3d} Arakawa K, Yachie N, Tomita M.(2007) E-Cell Simulation Environment 3D: 3-dimensional visualization of cellular simulation results.
ISMB 2007. http://ecell3d.iab.keio.ac.jp/

\bibitem{cyanobaci} Itaya M, Tsuge K, Koizumi M, Fujita K.(2005)
Combining two genomes in one cell: stable cloning of the Synechocystis PCC6803 genome in the Bacillus subtilis 168 genome.
Proc Natl Acad Sci U S A. 102(44):15971-6.

\bibitem{artificial_genome3} Gibson DG et al.(2010) Creation of a bacterial cell controlled by a chemically synthesized genome.
Science 329(5987):52-6

\bibitem{artificial_genome_rev_jpn3} 板谷光泰(2010) 「ゲノム合成の起承転結」科学 80(7):734-736

\end{thebibliography} 
