\documentstyle[12pt,epsf]{report}
\topmargin=-45pt
\oddsidemargin=0cm
\evensidemargin=0cm
\textheight=23.7cm
\textwidth=16cm

\def\thebibliography#1{\chapter*{References\markboth
 {References}{}}\list
 {[\arabic{enumi}]}{\settowidth\labelwidth{[#1]}\leftmargin\labelwidth
 \advance\leftmargin\labelsep
 \usecounter{enumi}}
 \def\newblock{\hskip .11em plus .33em minus .07em}
 \sloppy\clubpenalty4000\widowpenalty4000
 \sfcode`\.=1000\relax}
\let\endthebibliography=\endlist

% \def\thebibliography#1{\section*{References\markboth
%  {References}{}}\list
%  {[\arabic{enumi}]}{\settowidth\labelwidth{[#1]}\leftmargin\labelwidth
%  \advance\leftmargin\labelsep
%  \usecounter{enumi}}
%  \def\newblock{\hskip .11em plus .33em minus .07em}
%  \sloppy
%  \sfcode`\.=1000\relax}
% \let\endthebibliography=\endlist

\title{Computer Analyses of\\ Translation Initiation Sites of Genes}
\author{
 Tomita Laboratory ATG Project\\
 \begin{tabular}{|c|c|l|}
 \hline
  Name            & Student ID & Affiliation \\
 \hline
  Rintaro Saito   & 89531794 & Graduate School of Media and Governance \\
  Hidekazu Sasaki & 79401975 & Faculty of Policy Management \\
  Yuko Osada      & 79450924 & Faculty of Environmental Information \\
  Yukari Shimizu  & 79452137 & Faculty of Environmental Information \\
 \hline
\end{tabular}\\
% Takefuji Laboratory Neural Computing Group\\
Keio University Shonan Fujisawa Campus
}
\date{}
\begin{document}
\maketitle


\begin{Large}
\verb+       +Abstract of Master's Thesis

\vspace{1ex}
\verb+                       +Academic Year 1996\\
\end{Large}

\begin{LARGE}
\begin{center}
Computer Analyses of\\ Translation Initiation Sites of Genes
\end{center}
\end{LARGE}

\vspace{4ex}
\begin{small}
Gene is one of the main factors that determine how the organism will
be like. The information of gene is written on the molecule called DNA
in the cell. But organisms use only part of DNA as genes. The exact
mechanism of distinguishing gene regions in DNA is still unknown.
Previous research showed that there is some remarkable patterns in the 
boundary region of genes on DNA. 
In this paper, we focused on one of the boundaries called translation
initiation site. 
By introducing computer programs that we developed, comprehensive analyses
of sequencing results 
 in GenBank database(database for genetic sequences) were conducted.
This analyzing method is an original and new one.
We have discovered tendencies that would be important. For
example, there  
is general tendency to have fewer AUG trinucleotides around start
codons. And higher organisms tend to keep the upstream AUG trinucleotides
farther from start codons.  \\ 

% for translation  initiation.
% Translation is a process that ribosomes synthesize protein by scanning
% a mRNA from left to right.  Although ribosome initiates translation at
% the first found AUG trinucleotide in eucaryotes most of the time,
% it sometimes skips
% and ignores the first AUG and initiates translation at the second or
% third AUG trinucleotide.  On the other hand in procaryotes, 
% Shine-Dalgarno sequence
% seems to play a important role for translation initiation, but this
% sequence is not completely conserved. And there are many regions that have
% sequences very similar to Shine-Dalgarno sequences and does not serve as
% translation initiation site. 
% After all, the exact mechanism 
% of selecting AUG for translation initiation is not known.  
% By using GenBank database(database for nucleotide sequences) and 
% computer programs, comprehensive analyses were conducted.
% We have discovered tendencies that may be important.
% for translation  initiation.
% Some results show the new aspects of characteristics of translation 
% initiation and its site.
\end{small}

\begin{small}
This work was done by Tomita Laboratory ATG Projects.
Main roles of each members are briefly described below. 

\begin{description}
\item[Academic adviser:] M.Tomita
\item[Project leader:] R.Saito
\item[Survey:] R.Saito
\item[Construction of main analysis program] R.Saito
\item[Analyses of profile around start codons] 
H.Sasaki(program and analyses of
eucaryotes), \\Y.Osada(procaryotes), Y.Shimizu(mitochondria), R.Saito(adviser) 
\item[Analyses of entropy values around start codons] Y.Osada
\item[Analyses of GC/AT contents in 5'UTR] R.Saito
\item[Analyses of frequencies of AUG trinucleotides around start
codons] R.Saito
\item[Analyses of rate of leaky scanning and distance between two
AUGs] R.Saito,H.Sasaki
\item[Analyses of tendencies when two AUGs are close to each other]
R.Saito, H.Sasaki
\item[Analyses of relationship between initiation mechanism and
evolution]
R.Saito
\item[Analyses of frequencies of stop codons located downstream of
skipped AUGs] H.Sasaki

\end{description}
\end{small}

\vspace{2ex}
\noindent
Key Words
\begin{small}
\begin{center}
1.Gene\ \ \  2.Translation Initiation \ \ \ 3.Computer Analyses \ \ \ 4.Start Codon\ \ \ 5.Leaky Scanning

\end{center}
\end{small}
\vspace{2ex}

\begin{center}
{\large Keio University Graduate School of Media and Governance}
\end{center}

\begin{large}
\verb+                                             +Rintaro Saito
\end{large}

\tableofcontents


\chapter{Overview}
\label{overv}
\input{mp3_intr.tex}

\chapter{Materials and Methods}

In this chapter, original method used for the analyses is described.

\label{matmeth}
\input{mp11_ovrv.tex}


\chapter{Background}
\label{backgr}
\input{mp8_fdation.tex}

% \input{mp4_back.tex}

\section{Advance in the Research of Translation Initiation}
\label{surv}
\input{mp1_sur.tex}




\chapter{Result of This Research}

We conducted comprehensive computer analyses on translation initiation 
site by using our method described in chapter \ref{matmeth}.
In section \ref{prof_ent}, we conduct analyses of nucleotide
distributions around start codons to verify the results of previous
research and to discover new tendencies.
In section \ref{lowfreq}, we conduct comprehensive analyses from
various view points to examine our hypothesis that  AUG trinucleotides
near start codons confuse ribosomes. In section \ref{reinitia}, we
focus on stop codons located downstream of skipped AUGs to investigate the
significance of those codons for leaky scannings.

\label{resu}
\section{Profile and Entropy around Start Codons}
\label{prof_ent}

The purpose of this section is to analyze tendencies of nucleotide
distributions around start codons to find whether the tendencies are
consistent with previous research and to discover remarkable tendencies.
In previous study, Kozak conducted similar analyses with
699 vertebrate mRNA sequences\cite{label3}.
But our sequence data were taken from
GenBank database, which contains more than ten thousands mRNA sequences.



\input{mp2_prof.tex}

\section{On Low Frequencies of AUG Trinucleotides in front of Start Codons}
\label{lowfreq}

In this section, we focus on the skipped AUGs that are close to start
codons.
Ribosomes ignore AUGs that should not serve as start codons by leaky 
scanning and they select appropriate AUGs as start codons. But do
ribosomes confuse to select appropriate AUG as start codon if two AUGs
are close to each other? 
% Some experiments showed that AUG
% trinucleotide just in front of start codon will disturb translation
% initiation from the appropriate start codon in some organisms such as 
% yeast $Saccharomyces cerevisiae$\cite{label30}.
 We investigated this hypothesis from
4 approaches. That is frequencies of AUG trinucleotides around start codons,
rate of leaky scanning when two AUGs are close to each other,
consensus sequences when two AUGs are close, and relationship between
translation initiation mechanisms and evolutions.


\subsection{Frequencies of AUG trinucleotides around start codons}
 
\input{mp7_atgfreq.tex}

\subsection{Rate of leaky scanning and distance between two AUGs}

\input{mp12_rlk.tex}

\subsection{Tendencies when two AUGs are close to each other}

\input{mp9_atgnatg.tex}


\subsection{Relationship between initiation mechanism and evolution}

\input{mp6_avdist.tex}

\subsection{Discussion of the cases where two AUGs are close}

From our results, we suggest that there is evolutional pressure to 
eliminate AUG trinucleotides near start codons, presumably because AUG 
trinucleotides near start codons may confuse ribosomes. This
hypothesis is consistent with recent experimental research on yeast
$Saccharomyces\:cerevisiae$\cite{label30}.
The tendency to keep the AUG trinucleotides far from start codons
is
weaker in lower organisms, presumably because lower organisms have
more reliable mechanisms to identify the appropriate AUG as a start codon.

However the existing AUGs, which are close to each other, does not
seem to have much influence on the fidelity of translation initiation. 
%Thus the
%hypothesis that upstream AUGs near start codons are
%unpreferable holds in the long period of time. 
But in cases where two
AUGs are separated by two bases, strong consensus sequences may be 
needed to avoid leaky scannings.



\section{Frequencies of Stop Codons Located Downstream of Skipped AUGs}
\label{reinitia}
\input{mp10_stop.tex}


\newpage
% \setcounter{page}{30}
\chapter{Conclusion}
\label{concl}
\input{mp5_concl.tex}

\chapter*{Acknowledgements}
The author thanks to associate professor Masaru Tomita for giving me
suggestions and comments. And thanks to professor Takemochi Ishii and
associate professor Yoshiyasu Takefuji for giving me a comment on this
paper. And thanks to Yoshimi Toda and Junko Hara for checking my
paper. And thanks to my project members, Hidekazu Sasaki, Yuko Osada,
and Yukari Shimizu for working with me in the project.

\newpage
\begin{thebibliography}{99}
% \addcontentsline{toc}{chapter}{References}
\bibitem{label0} Bruce Alberts et al.,
Molecular Biology of THE CELL third edition,
Garland Publishing Inc.

\bibitem{label3} Kozak,M.(1987) An analysis of 5'-noncoding sequences from
699 vertebrate messenger RNAs,Nucleic Acids Research Volume 15 pp.8125-8148

\bibitem{label1} Kozak,M.(1995),Adherence to the first-AUG rule when a 
second AUG codon follows closely upon the first,
Proc.Natl.Acad.Sci.USA Vol.92 pp.2662-2666

\bibitem{label2} Kozak,M.(1990),Downstream secondary structure facilitates
recognition of initiator codons by eukaryotic ribosomes,
Proc.Natl.Acad.Sci.USA VOl.87.pp.8301-8305

\bibitem{label4} Kozak,M.(1989),The Scanning Model for Translation:An Update,
J.Cell.Biol 108,pp.229-241

\bibitem{label6} Kozak,M.(1989),Context Effects and Inefficient Initiation
 at Non-AUG Codons in Eucaryotic Cell-Free Translation Systems,

\bibitem{label12} Kozak,M.(1992),Regulation of Translation in Eucaryotic
Systems,Annuu.Rev.Cell Biol.8:197-225

\bibitem{label5} Slunsher, L. et al(1991),mRNA leader length and initiation
codon context determine alternative AUG selection for the yeast gene MOD5,
Proc.Natl.Acad.Sci.USA Vol.88,pp.9789-9793


\bibitem{label8} Imataka,H. et al(1994), Cell-specific Translational Control
of Transcription Factor BTEB Expression,J.Biol.Chem. Vol.269 pp.20668-20673

\bibitem{label7} Shine,J. and Dalgarno,L.(1974). The 3'-terminal sequence
of Escherichia coli 16S ribosomal RNA:Complementarity to nonsense triplets
and ribosome binding sites. Proc.Nat.Acad.Sci.USA Vol.71 pp1342-1346

\bibitem{label9} Hongyun Chen, et al(1994),
Determination of the optimal aligned spacing between the Shine-Dalgarno
sequence and the translational initiation codon of Escherichia coli mRNAs,
Nuc.Acid.Res. Vol.22 pp.4953-4957

\bibitem{label10} Rinke-Appel,J. et al(1994),
 Contacts between 16S ribosomal RNA and
mRNA, within the spacer region separating the AUG initiator codon and the
Shine-Dalgarno sequence; a site directed cross-linking study,
Nuc.Acid.Res. Vol.22 pp.3018-3025

\bibitem{label11} Schneider,T et al.(1986),
Information Content of Binding Sites on Nucleotide Sequences,
J.Mol.Biol.188,pp.415-431

\bibitem{label18} Kozak,M.(1987),
 Effects of intercistronic length on the efficiency of reinitiation
by eucaryotic ribosomes,
Mol.Cell.Biol.10,pp.3438-3445

\bibitem{label19} Kozak,M.(1991),
 A short leader sequence impairs the fidelity of initiation by eukaryotic
ribosomes, Gene Expr, 2, pp.111-115

\bibitem{label20} Claudio O.Gualerzi and Cynthia L.Pon(1990),
Initiation of mRNA Translation in Procaryotes,
Biochem.Vol.29,pp.5881-5889

\bibitem{label21} Kozak,M.(1986),
Influences of mRNA secondary structure on initiation by eucaryotic 
ribosomes,
Proc.Natl.Acad.Sci.83,pp.2850-2854


\bibitem{label22} Kozak,M.(1983)
Comparison of initiation of protein synthesis in procaryotes, eucaryotes,
and organelles,
Microbiol Rev Vol.47 pp.1-45

\bibitem{label23} IG.Ivanov, et al.(1995)
A second putative mRNA binding site on the Escherichia coli ribosome.
Gene 160:75-79

\bibitem{label24} Hannig,EM.(1995)
Protein synthesis in eucaryotic organisms: new insights into the function
of translation initiation factor eIF3.
Bioessays 17:915-919

\bibitem{label25} Herman, A.DE BOER et al.(1990)
Sequences within Ribosome Binding Site Affecting Messenger RNA Translatability
and Methods to Direct Ribosomes to Single Messenger RNA Species.
Methods Enzymol. 185:103-114

\bibitem{label26} Maarten H. de Smit. et al.(1994)
Translational Initiation on Structured Messengers.
Another Role for the Shine-Dalgarno Interaction.
J.Mol.Biol. 235:173-184

\bibitem{label27} Claire M.Fraser et al.(1995)
The Minimal Gene Complement of Mycoplasma genitalium.
Science 270:397-403

\bibitem{label28}
Barrick D. et al.(1994)
Quantitative analysis of ribosome binding sites in E.coli.
Nuc.Acid.Res. 22:1287-1295

\bibitem{label29}
Mehdi H. et al.(1990)
Initiation of translation at CUG, GUG and ACG codons in mammalian
cells.
Gene 91:173-178

\bibitem{label30}
Ding-Fang Yun et al.(1996)
mRNA sequences influencing translation and the selection of AUG
initiator codons in the yeast Saccharomyces cerevisiae.
Mol.Microbiol.19:1225-1239

\bibitem{label31}
Luukkonen BG et al.(1995)
Efficiency of reinitiation of translation on human immunodeficiency
virus type 1 mRNAs is determined by the length of the upstream open
reading frame and by intercistronic distance.
J.Virol. 69:4086-4094

\end{thebibliography}

\end{document}




