\documentstyle[12pt,epsf]{report}



\begin{document}
\section{On Low Frequencies of AUG Trinucleotides in front of Start Codons}
In this section, we focus on the skipped AUGs that are close to start
codons.

Ribosomes ignore AUGs that should not serve as start codons by leaky 
scanning and they select appropriate AUGs as start codons. But do
ribosomes confuse to select appropriate AUG as start codon if two AUGs
are close to each other? We investigated this hypothesis from
4 approaches. That is frequencies of AUG trinucleotides around start codons,
rate of leaky scanning when two AUGs are close to each other,
consensus sequences when two AUGs are close, and relationship between
translation initiation mechanisms and evolutions.


\subsection{Frequencies of AUG trinucleotides around Start Codons}
 
\input{mp7_atgfreq.tex}

\subsection{Rate of Leaky Scanning and Distance between 2 AUGs}

\input{mp12_rlk.tex}

\subsection{Tendencies When Two AUGs are Close to Each Other}

\input{mp9_atgnatg.tex}


\subsection{Relationship between initiation mechanism and evolution}

\input{mp6_avdist.tex}

\subsection{Discussion of the cases where two AUGs are close}

From our results, we suggest that there is evolutional pressure to 
eliminate AUG trinucleotides near start codons. This pressure is 
lower in lower organisms, presumably because lower organisms have
more reliable mechanisms to identify the appropriate AUG as a start codon.

However the existing AUGs, which are close to each other, does not
seem to have much influence on translation initiation. Thus the
hypothesis that upstream AUGs near start codons are
unpreferable holds in the long period of time. But in cases where two
AUGs are separated by two bases, strong consensus sequences may be 
needed to avoid leaky scannings.

\end{document}
