\section{Recent Advance in This Field}

Although bio-informatics is a new field of research, 
there are many field of research.

\subsection{Sequencing}

Sequencing is the reading of the nucleic acid or amino acid sequences.
By using method called PCR, many DNA sequence data are read.
There are some organisms whose whole genome data were read.
That are, for example, bacteria like {\it Mycoplasma Genitalium},
{\it Haemophilus influenzae}. Whole human genome will be read in
Human Genome Project which is in progress by many countries.
And whole {\it E.Coli} genome data will also be read.
Those data will be open to the public through
the internet. To determine where the genes are in the sequence,
biological experiments must be done. Gene finding algorithm may be
helpful if it can appropriately identify where the genes are, but
the prediction rate is about 90\%. It needs improvement.

\subsection{Signal Analyses}

In DNA sequences, there are many signals such as transcription start signal,
transcription termination signal, etc.


\subsection{Prediction of the function of proteins}

In bio-informatics,prediction is done by homology search.

\subsection{Investigation of evolutions from the molecular level}

By this research, we can know how the species evolved in molecular level
rather than fossil level. This can be improved by alignment algorithms.


\subsection{Prediction of secondary or tertialy structure of nucleic 
acid or protein}

RNA secondary structure prediction is done by Zuker.


\subsection{Simulation of biological systems}

Methabolic simulation is done by Peter Karp, Douglas Brutlag.

\section{Ultimate Goal of This Field}

Ultimate goal of this field is to completely understand life from 
biological data such as DNA sequences. If this goal is accomplished, then
we can cure disease by modifying DNA sequences. And we can change our body
to whatever we wish in the same way. Those are the contribution to medicine.
Also in the same way, we can produce many kinds of foods. That is 
the contribution to biotechnology.

But we always have to keep in mind what we should not do with those 
technologies.




