\begin{thebibliography}{99}
\bibitem{ent_TATA} 田村隆明, 新転写制御のメカニズム, 羊土社, 2000
\bibitem{ent_label_SD} 
Shine J, Dalgarno L,
The 3'-terminal sequence of Escherichia coli 16S ribosomal RNA: complementarity to nonsense triplets and ribosome binding sites,
Proc Natl Acad Sci U S A. 71(4):1342-6, 1974
\bibitem{Osada1} Osada Y, Saito R, Tomita M,
Analysis of base-pairing potentials between 16S rRNA and 5' UTR for 
translation initiation in various prokaryotes,
Bioinformatics, 15(7-8):578-81, 1999
\bibitem{ent_cyano_SD} Sazuka T, Ohara O.,
Sequence features surrounding the translation initiation sites assigned 
on the genome sequence of Synechocystis sp. strain PCC6803 by 
amino-terminal protein sequencing, DNA Res. 3(4):225-32, 1996
\bibitem{info_theor1} 島田良作, わかる情報理論, 日新出版, 1982
\bibitem{integra1} 小田原宏行, 微分積分学入門, 北樹出版, 1986
\bibitem{info_cont} Schneider TD, Stormo GD, Gold L, Ehrenfeucht A,
Information content of binding sites on nucleotide sequences,
J Mol Biol. 188(3):415-31, 1986 
\bibitem{statistics1} 稲垣宣生, 数理統計学, 裳華房, 1990
\bibitem{texbook}
Mount DW著、岡崎康司、坊農秀雅監訳，バイオインフォマティクス，メディカルサイエンスインターナショナル，2002.
\bibitem{align_SW}
Smith TF, Waterman MS,
Identification of common molecular subsequences,
J Mol Biol. 147(1):195-7,1981 
\bibitem{Bioinfo_eddy} Durbin, R, Eddy, SR, Krogh, A, Mitchison, G著,
阿久津達也, 浅井潔, 矢田哲士訳, バイオインフォマティクス-確率モデルによる
遺伝子配列解析-, 医学出版, 2001
\bibitem{ncRNA_numassan} Numata K, Kanai A, Saito R, Kondo S, Adachi J, Wilming LG, Hume DA, 
Hayashizaki Y, Tomita M; RIKEN GER Group; GSL Members,
Identification of putative noncoding RNAs among the RIKEN mouse 
full-length cDNA collection, Genome Res. 13(6B):1301-6, 2003  
\bibitem{Bioinfo_endo} Setubal, JC, Meidanis, J著, 五條堀孝監訳, 遠藤俊徳訳,
分子生物学のためのバイオインフォマティクス入門, 共立出版, 2001
\bibitem{biophysA} 白井道雄, 入門物理化学, 実教出版, 1978
\bibitem{zuker} Zuker M, Stiegler P,
Optimal computer folding of large RNA sequences using thermodynamics and auxiliary information,
Nucleic Acids Res. 9(1):133-48, 1981 
\bibitem{biotech_computer} 中村春木, 中井謙太, バイオテクノロジーのための
コンピュータ入門, コロナ社, 1995

\bibitem{label_Derisi} DeRisi JL, Iyer VR, Brown PO,
Exploring the metabolic and genetic control of gene expression on a genomic scale,
Science 278(5338):680-686, 1997 
\bibitem{label_scer_marray} Hughes TR, Marton MJ, Jones AR, Roberts CJ, Stoughton R, Armour CD, Bennett HA, Coffey E, Dai H, He YD, Kidd MJ, King AM, Meyer MR, Slade D, Lum PY, Stepaniants SB, Shoemaker DD, Gachotte D, Chakraburtty K, Simon J, Bard M, Friend SH,
Functional discovery via a compendium of expression profiles,
Cell, 102(1):109-26,2000

\bibitem{label_Eisen}
Eisen MB, Spellman PT, Brown PO, Botstein D,
Cluster analysis and display of genome-wide expression patterns,
Proc Natl Acad Sci U S A. 95(25):14863-14868, 1998
\bibitem{MA_tanoshi} 藤沢偉作, 楽しく学べる多変量解析,  
現代数学社, 1985

\bibitem{multi_nonmat} 田中豊, 脇本和昌著, 多変量統計解析法, 現代数学社, 1983

\bibitem{machine_L} Mitchell, TM, Machine Learning, WBC/McGraw-Hill, 1997

\bibitem{cbias_mut_trans1} 
Sharp PM, Li WH,
An evolutionary perspective on synonymous codon usage in unicellular organisms,
J Mol Evol. 24(1-2):28-38,1986
\bibitem{cbias_mut_trans2} 
Shields DC, Sharp PM,
Synonymous codon usage in {\it Bacillus subtilis} reflects both translational selection and mutational biases,
Nucleic Acids Res. 15(19):8023-40,1987 
\bibitem{label_CAI} Sharp PM, Li WH, 
The codon Adaptation Index--a measure of directional synonymous codon 
usage bias, and its potential applications,
Nucleic Acids Res. 15(3):1281-95,1987

\bibitem{linear_alge1} 吉原健一,寺田敏司, 現代線形代数学入門, 学術図書出版社, 1991
\bibitem{R_Tips} 舟尾暢男, The R Tips データ解析環境Rの基本技・グラフィックス活用術, 九天社, 2005

\bibitem{label_codon_opinion1}
Duret L,
Evolution of synonymous codon usage in metazoans,
Curr Opin Genet Dev. 12(6):640-649, 2002
\bibitem{label_codon_GC}
Muto A, Osawa S,
The guanine and cytosine content of genomic DNA and bacterial evolution,
Proc Natl Acad Sci U S A. 84(1):166-169,1987
\bibitem{miss_CA} Perriere, G., Thioulouse, J.
Use and misuse of correspondence analysis in codon usage studies.
Nucleic Acids Res. 30(20):4548-55,2002

\bibitem{CA_intro} Weller SC, Romney AK, Metric Scaling Correspondence Analysis,
SAGE Publications Inc., 1990

\bibitem{NN_arch} デイホフ J著, 桂井浩訳, 
ニューラルネットワークアーキテクチャー入門, 森北出版, 1992
\bibitem{NN_comp} ビール R, ジャクソン T著, 八名和夫監訳, 
ニューラルコンピューティング入門, 海文堂出版, 1993

\bibitem{label_stats_1} 柴田文明, 理工系の基礎数学７ 確率・統計, 岩波書店, 1996
\bibitem{label_stats_2} 小針あき宏, 確率・統計入門 岩波書店, 1973
\bibitem{econostats} 岩田暁一, 経営分析のための統計的方法, 東洋経済新報社, 1983

\bibitem{label1896} Basharin, G.P., On a statistical estimate for
	the entropy of a sequence of independent random
	variables. Theory Probability Appl. 4: 333-336, 1959

\bibitem{label5050} Wilks, SS, Mathematical Statistics, John Wiley,
\verb+&+ Sons, 1962
\bibitem{statistics_v1} ウィルクス SS著, 田中英之, 岩本誠一訳,
数理統計学１, 東京図書出版, 1972
\bibitem{statistics_v2} ウィルクス SS著, 田中英之, 岩本誠一訳,
数理統計学２, 東京図書出版, 1972
\bibitem{integra11} 笠原晧司, 微分積分学, サイエンス社, 1985
\bibitem{statistics2} 東京大学教養学部統計学教室編, 自然科学の統計学, 東京大学出版会, 1992

\end{thebibliography}

\endinput
