\documentstyle[doublespace,11pt,epsf]{report}
\topmargin=-45pt
\oddsidemargin=0cm
\evensidemargin=0cm
\textheight=23.7cm
\textwidth=16cm

\def\thebibliography#1{\chapter*{References\markboth
 {References}{}}\list
 {[\arabic{enumi}]}{\settowidth\labelwidth{[#1]}\leftmargin\labelwidth
 \advance\leftmargin\labelsep
 \usecounter{enumi}}
 \def\newblock{\hskip .11em plus .33em minus .07em}
 \sloppy\clubpenalty4000\widowpenalty4000
 \sfcode`\.=1000\relax}
\let\endthebibliography=\endlist

% \def\thebibliography#1{\section*{References\markboth
%  {References}{}}\list
%  {[\arabic{enumi}]}{\settowidth\labelwidth{[#1]}\leftmargin\labelwidth
%  \advance\leftmargin\labelsep
%  \usecounter{enumi}}
%  \def\newblock{\hskip .11em plus .33em minus .07em}
%  \sloppy
%  \sfcode`\.=1000\relax}
% \let\endthebibliography=\endlist


\title{Computer Analyses of Translation Initiation Sites of Genes}

\author{Rintaro Saito\\
Laboratory for Bioinformatics(Tomita's Laboratory)\\
Graduate School of Media and Governance\\
Keio University SFC %\\
%\verb+ +\\
%\verb+ +\\
%\verb+ +\\
%\begin{tabular}{|c|}
%\hline
%{\LARGE 試作版}\\
%\hline
%\end{tabular}%\\
%\verb+ +\\
%doublespaceのフォーマットで書かれている場合、\\
%実際のページ数より40\%弱程度多くなります。
}


\begin{document}
\maketitle

\begin{abstract}
Translation initiation is a process by which a ribosome initiates the
translation on mRNA. Although it is well studied, many questions about
its mechanisms still remain unanswered. In this work, translation
initiation sites were analyzed by the computer from various points of
view. Following four points are the main results. (a) Most consensus
sequences that appear around start codons are consistent with the
hypotheses of Shine-Dalgarno and Kozak. However, there are some
exceptions in some species such as {\it M.genitalium}. Those species
seem to use alternative mechanisms to initiate
translation. Shine-Dalgarno sequences may be required even when
ribosomes can reinitiate the translations.  (b) In some procaryotes,
especially in {\it E.coli}, frequencies of Shine-Dalgarno sequences are
fewer when start codons are GTG. {\it E.coli} seem to increase
translation level of highly expressed genes by using both strong
Shine-Dalgarno sequences and strong start codons(ATG). (c) In most
procaryotes, there are strong base correlations between adjacent bases
upstream of start codons, presumably to form efficient consensus
sequences. However, this tendency was not clearly observed in
eucaryotes, presumably because 18S rRNA does not bind to consensus
sequences on mRNA. In {\it H.sapiens} and {\it M.musclus}, dinucleotides
immediately before start codons tend to consist of the same bases.  (d)
Frequency of ATG triplets decrease around start codons in most
organisms.  And how they decrease seems to correspond to how ribosomes
recognize translation initiation sites, i.e., eubacterial ribosomes bind
directly to translation initiation sites whereas eucaryotic ribosomes
scan mRNA to find start codons. Similar analyses of frequency of ATG
triplets in archaebacteria suggest that in some of them, ribosomes scan
mRNA to bind to Shine-Dalgarno sequences.

\end{abstract}

\begin{abstract}

%\begin{center}
%\begin{bf}
%{\large 概要}
%\end{bf}
%\end{center}

翻訳開始とはmRNAと呼ばれる細胞内高分子がタンパク質に変換される最初のプロ
セスのことである。しかしそのプロセスに関しては不明な点が多く、mRNA上のど
のような配列とどのような因子が翻訳開始に関与しているのか、まだ完全には解
明されていない。そこで様々な生物のmRNAまたはそのもととなるDNAの配列を様々
な角度からコンピュータで分析し、配列中に隠れた傾向の抽出を行った。

その結果、以下のようなことが判明した。(a) 翻訳開始領域に出現するコンセン
サス配列はほとんどの場合、Shine-Dalgarnoの仮説やKozakの仮説に基づく配列
に一致する。しかし、{\it M.genitalium}など一部の種ではShine-Dalgarno配列
以外の配列パターンが発見された。Shine-Dalgarno配列は翻訳の再開
(reinitiation)の時にもおおむね必要であることが判明した。(b) {\it E.coli}
では、開始コドンがGTGになるとShine-Dalgarno配列のコンセンサスが弱くなる。
従ってこの種では開始コドンとShine-Dalgarno配列両方のコンセンサスを強くす
ることによって発現量の高い遺伝子の翻訳開始の効率を上げていると思われる。
(c)原核生物ではリボソーム結合部位付近にコンセンサス配列を作りやすくする
ような塩基間の相互作用があることが判明した。しかしこれは真核生物に関して
は必ずしもあてはまらなかった。これはおそらく真核生物では翻訳開始時に18S
rRNAがmRNA上のコンセンサス配列に結合するということが一般的にはないためと
思われる。{\it H.sapiens}や{\it M.musclus}には開始コドン直前の2塩基を同
じものにするような相互作用があることがわかった。(d)ATGという配列パターン
が開始コドン周辺ではリボソームの適切な翻訳開始反応を阻害する可能性がある
ことがコンピュータ解析により確かめられた。またATGの頻度の分布を様々な生
物で比較し、原核生物ではリボソームが直接翻訳開始領域に結合するのに対し、
真核生物ではリボソームがmRNA上を移動して翻訳開始領域を探すという従来の
学説をサポートする結果を得ている。さらに古細菌と呼ばれる生物種は原核生物
と真核生物両方の翻訳開始のメカニズムを使っている可能性があることが判明し
た。

\end{abstract}


\tableofcontents

\chapter{Introduction}

\input{dp1_intr.tex}

\chapter{Translation Initiation}
\label{survey}

\input{dp2_survey.tex}

\chapter{Consensus Patterns}
\label{consensus}

\input{consensus.tex}

\chapter{Start Codons and Shine-Dalgarno sequences}
\label{chiyo}

\input{chiyo3.tex}

\chapter{Base Correlations}
\label{correlation}

\input{correlation.tex}

\chapter{Scanning Models}
\label{neg_atg}

\input{neg_atg.tex}

\chapter{Conclusion}
\label{conclusion}

\input{dp5_concl.tex}


\chapter*{Acknowledgement}

First of all, I would like to thank to Professor Masaru Tomita, my
thesis advisor, Professor Nobuo Saito, Professor Koichi Furukawa,
Professor Sadakazu Aiso, and Professor Yasufumi Murakami. And I also
would like to thank Yuko Osada, Yosuke Ozawa, Sadahiro Kumagai, Naoki
Kuzuno, Satoshi Hanaoka, Minoru Takayama, Hidekazu Sasaki, Yukari
Shimizu, Chiyo Yasui, Chiaki Yajima and Tatsuo Matsumto for working with me
in this field. I would also like to thank Tomonobu Ozaki for giving me
much advise on computational issues.



\begin{thebibliography}{999}
% \addcontentsline{toc}{chapter}{References}

\bibitem{label891} Adhin, M.R. and van Duin, J.(1990) Scanning model for translational reinitiation in eubacteria. J.Mol.Biol. 213:811-818

\bibitem{label0} Alberts,B., Bray, D., Lewis, J., Raff, M., Roberts,
	K. and Watson, J.D.(1994) Molecular Biology of THE CELL third
	edition,Garland Publishing Inc.

\bibitem{label7009} Andersson, S.G., Zomorodipour, A., Andersson, J.O.,
	Sicheritz-Ponten, T., Alsmark, U.C., Podowski, R.M., Naslund,
	A.K., Eriksson, A.S., Winkler, H.H. and Kurland, C.G.(1998) The
	genome sequence of {\it Rickettsia prowazekii} and the origin of
	mitochondria. Nature 396:133-143

\bibitem{label615} Arimura, H., Fujino, R., Shinohara, T. and
Arikawa, S.(1994) Protein motif discovery from positive examples by
minimal multiple generalization over regular patterns. 
Genome Informatics 1994 39-48

\bibitem{label898} Atkins, J.F.(1979) Is UAA or UGA part of the
recognition signal for ribosomal initiation? Nuc.Acid.Res. 7:1035-1041

\bibitem{arclabel10} Band, L. and Henner, D.J.(1984) {\it Bacillus subtilis} requires a``stringent'' Shine-Dalgarno region for gene expression. DNA 3:17-21

\bibitem{label28} Barrick, D., Villanueba, K., Childs, J., Kalil, R.,
	Schneider, T.D., Lawrence, C.E., Gold, L. and Stormo,
	G.D.(1994) Quantitative analysis of ribosome binding sites in
	{\it E.coli}. Nuc.Acid.Res. 22:1287-1295

\bibitem{label1896} Basharin, G.P.(1959) On a statistical estimate for
	the entropy of a sequence of independent random
	variables. Theory Probability Appl. 4: 333-336

\bibitem{label2020} Bell, S.D. and Jackson, S.P.(1998) Transcription and
translation in Archaea: a mosaic of eukaryal and bacterial
features. Trends Microbiol. 6:222-228 

\bibitem{label1700} Benson, D.A., Boguski, M.S., Lipman, D.J., Ostell,
	J., Ouellette, B.F., Rapp, B.A. and Wheeler, D.L.(1999)
	GenBank. Nuc.Acid.Res. 27:12-17

\bibitem{label1307} Berkhout, B., Kastelein, R.A. and van Duin, J.(1985)
	Translational interference at overlapping reading frames in
	prokaryotic messenger RNA. Gene 37:171-179 

\bibitem{label810} Bisant D. and Maizel, J.(1995) Identification of
	ribosome binding sites in {\it Escherichia coli} using neural
	network models. Nuc.Acid.Res. 23:1632-1639

\bibitem{label1065} Blattner, F.R., Plunkett III, G., Bloch C.A., Perna,
	N.T., Burland V., Riley, M., Collado-Vides, J., Glasner, J.D.,
	Rode, C.K., Mayhew, G.F., Gregor, J., Davis, N.W., Kirkpatrick,
	H.A., Goeden, M.A., Rose, D.J., Mau, B. and Shao, Y.(1997) The complete genome sequence fo {\it Escherichia coli} K-12. Science 277:1453-1474


\bibitem{label993} Brodovsky M. and McIninch, J.(1993) GENMARK: Parallel gene recognition for both DNA strands. Computers Chem. 17:123-133

\bibitem{label635} Bucklew, J.A.(1990) Large Deviation Techniques in Decision, Simulation, and Estimation. John Wiley and Sons, Inc., New York.

\bibitem{label2080} 
Bult, C.J., White, O., Olsen, G.J., Zhou, L., Fleischmann, R.D.,
Sutton. G.G., Blake, J.A., FitzGerald, L.M., Clayton, R.A., Gocayne,
J.D., Kerlavage, A.R., Dougherty, B.A., Tomb, J.F., Adams, M.D.,
Reich, C.I., Overbeek, R., Kirkness, E.F., Weinstock, K.G., Merrick,
J.M., Glodek, A., Scott, J.L., Geoghagen, N.S.M. and Venter, J.C.(1996)
Complete genome sequence of the
methanogenic archaeon, {\it Methanococcus jannaschii}. Science
273:1058-1073 

\bibitem{label9} Chen, H., Bjerknes, M., Kumar, R. and  Jay, E.(1994) Determination of the optimal
	aligned spacing between the Shine-Dalgarno sequence and the
	translational initiation codon of {\it Escherichia coli}
	mRNAs. Nuc.Acid.Res. 22:4953-4957

\bibitem{mglabel11} Dalphin, M.E., Brown, C.M., Stockwell, P.A. and
	Tate, W.P.(1998) The translational signal databese, TransTerm,
	is now a relational database. Nuc.Acid.Res., 26:335-337, 1998.

\bibitem{label25} de Boer, H.A. and Hui, A.S.(1990) Sequences within
	ribosome binding site affecting messenger RNA translatability
	and methods to direct ribosomes to single messenger RNA
        species. Methods Enzymol. 185:103-114

\bibitem{lsstr1} de Smit, M. and van Duin, J.(1994) Control of translation by mRNA secondary structure in {\it Escherichia coli}. A quantitative analysis of literature data. J.Mol.Biol. 244:144-150

\bibitem{label26} de Smit, M. and van Duin, J.(1994) Translational
	Initiation on Structured Messengers.Another Role for the
	Shine-Dalgarno Interaction. J.Mol.Biol. 235:173-184


\bibitem{label1915} Dominguez, D., Altmann, M., Benz, J., Baumann, U.
	and  Trachsel, H.(1999) Interaction of translation initiation
	factor eIF4G with eIF4A in the yeast {\it Saccharomyces cerevisiae}. 
        J.Biol.Chem. 274:26720-6 

\bibitem{label3001} Dunn, R., McCoy, J., Simsek, M., Majumdar, A.,
	Chang, S. H., RajBhandary, U. L. and Khorana, H.G.(1981) The
	bacteriaorhodopsin gene. Proc.Natl.Acad.Sci.USA 78:6744-6748  


\bibitem{mglabel10} Dybvig K. and Voelker, L.(1996) Molecular biology of
	mycoplasmas. Annu.Rev.Microbiol. 50:25-57

\bibitem{label1501} Etchegaray, J.P. and  Inouye, M.(1999) Translational
	enhancement by an element downstream of the initiation codon in
	{\it Escherichia coli}. J Biol Chem 274:10079-10085

\bibitem{mglabel14} Fraser, C.M., Gocayne, J.D. White, O., Adams, M.D.,
Clayton, R.A., Fleischmann, R.D., Bult, C.J., Kerlavage,
A.R., Sutton, G., Kelley, J.M. et al.(1995) The minimal gene
	complement of {\it Mycoplasma genitalium}. Science 270:397-403

\bibitem{label1005} Frishman, D., Mironov, A. and Gelfand, M.(1999)
Starts of bacterial genes: estimating the reliability of computer
predictions. Gene 234:257-265

\bibitem{label8001} Gold, L., Pribnow, D., Schneider, T., Shinedling,
	S., Singer, B.S. and Stormo, G.(1981) Translational initiation
	in prokaryotes. Annu.Rev.Microbiol. 35:365-403

\bibitem{label1920} Gray, N.K. and  Wickens, M.(1998) Control of translation
	initiation in animals. Annu.Rev.Cell Dev. Biol. 14:399-458   

\bibitem{label875} Grillo, G., Attimonelli, M., Liuni, S. and Pesole,
	G.(1996) CLEANUP: A fast computer program for removing
	redundancies from nucleotide sequence databases. CABIOS 12:1-8

\bibitem{label7219} Gr\"unert, S. and Jackson, R.J.(1994)
The immediate downstream codon strongly influences the
efficiency of utilization of eukaryotic translation initiation
codons. EMBO J 13:3618-30 

\bibitem{label20} Gualerzi,C.O., and Pon,C.L.(1990) Initiation of mRNA
	Translation in Procaryotes. Biochem. 29:5881-5889

\bibitem{label7008} Guig\'o, R., Knudsen, S., Drake, N. and Smith,
	T.(1992) Prediction of gene structure. J.Mol.Biol. 226:141-157


\bibitem{label7078} Hagenbuchle, O., Santer, M., Steitz, J.A. and
Mans, R.J.(1978) Conservation of the primary structure at the 3' end
of 18S rRNA from eucaryotic cells. Cell 13:551-563

\bibitem{label921} Hannenhalli, S.S., Hayes, W.S., Hatzigeorgiou, A.G. 
and Fickett. J.W.(1999) Bacterial start site prediction. Nuc.Acid.Res. 
27:3577-3582.

\bibitem{label24} Hannig, E.M.(1995) Protein synthesis in eucaryotic
	organisms: new insights into the function of translation
	initiation factor eIF3. Bioessays 17:915-919

\bibitem{label4003} Harms, E., Higgins, E., Chen, J.W. and Umbarger, H.E.(1988)
Translational coupling between the ilvD and ilvA genes of
{\it Escherichia coli}. J. Bacteriol. 170:4798-4807 

\bibitem{label4807} Hemmings-Mieszczak, M. and  Hohn, T.(1999)
A stable hairpin preceded by a short open reading frame promotes
nonlinear ribosome migration on a synthetic mRNA leader. RNA 5:1149-1157 

\bibitem{label1098} Herbert, A.(1996) RNA editing, introns and evolution. TIG 12:6-9

\bibitem{label625} Hertz, G.Z., Hartzell, W. III and Stormo. G.D.(1990)
Identification of consensus patterns in unaligned DNA sequences known
to be functionally related. Bioinformatics 6:81-92

\bibitem{label629} Hertz, G.Z. and Stormo, G.D.(1999)
     Identifying DNA and protein patterns with statistically significant
     alignments of multiple sequences.
     Bioinformatics. 15:563-577

\bibitem{label511} Herzel, H., Trifonov, E.N., Weiss, O., Gro$\beta$e, I.(1998) Interpreting correlations in biosequences. Physica A 249:449-459

\bibitem{label503} Himmelreich, R., Hilbert H., Plagens, H., Pirkl, E. Li, B. and Herrmann R.(1996) Complete sequence analysis of the genome of the bacterium {\it Mycoplasma pneumoniae}. Nuc.Acid.Res. 24:4420-4449

\bibitem{label7011} Hui, A., and de Boer, H.A.(1987)
Specialized ribosome system: preferential translation of a single
mRNA species by a subpopulation of mutated ribosomes in
{\it Escherichia coli}. Proc.Natl.Acad.Sci.USA 84:4762-4766

\bibitem{label23} Ivanov, I.G., Alexandrova, R.A., Dragulev, B.P. and
	AbouHaidar, M.G.(1995) A second putative mRNA binding site on
	the {\it Escherichia coli} ribosome. Gene 160:75-79

\bibitem{label8} Imataka, H., Nakayama, K., Yasumoto, K., Mizuno, A.,
	Fujii-Kuriyama, Y. and Hayami, M.(1994) Cell-specific translational
control of transcription factor BTEB
	expression. J.Biol.Chem. 269:20668-20673

\bibitem{label7333} Jacques, N. and Dreyfus, M.(1990) Translation
	initiation in {\it Escherichia coli}: old and new questions.
        Mol.Microbiol. 4:1063-1067

\bibitem{label701} Joshi, C.P., Zhou, H., Huang, X. and Chiang,
V.L.(1997) Context sequences of translation initiation codon in
plants. Plant Mol.Biol. 35:993-1001

\bibitem{label605} Jonassen, I., Collins, J.F. and Higgins,
D.G.(1995) Finding flexible patterns in unaligned protein
sequences. Protein Sci. 4:1587-1595

\bibitem{label1189} Karimi, R., Pavlov, M.Y., Buckingham, R.H. and
	Ehrenberg, M.(1999) Novel roles for classical factors at the
	interface between translation termination and initiation. 
        Mol. Cell 3:601-9 

\bibitem{label301} Karlin, S. and Cardon, L.R.(1994) Computational DNA
sequence analysis. Annu.Rev.Microbiol. 48:619-654

\bibitem{label2051} Keeling, P.J. and Doolittle, W.F.(1995) Archaea: Narrowing
the gap between prokaryotes and eukaryotes. Proc.Natl.Acad.Sci.USA 92:5761-5764

\bibitem{label2052} Keeling, P.J., Fast, N.M. and McFadden, G.I.(1998) Evolutionary relationship between translation initiation factor eIF-2$\gamma$ and
Selenocysteine-Specific elongation factor SELB: Change of function in
translation factors. J.Mol.Evol. 47:649-655

\bibitem{label1910} King, R.D., Muggleton, S.H., Srinivasan, A. and
	Sternberg, M.J.(1996) Structure-activity relationships derived
	by machine learning: the use of atoms and their bond
	connectivities to predict mutagenicity by inductive logic
	programming. Proc.Natl.Acad.Sci. U.S.A. 93:438-42 

\bibitem{label22} Kozak,M.(1983) Comparison of initiation of protein
	synthesis in procaryotes, eucaryotes, and organelles. Microbiol. Rev. 47:1-45

\bibitem{label1112} Kozak,M.(1984) Selection of initiation sites by eucaryotic ribosomes: effect of inserting AUG triplets upstream from the coding sequence for preproinsulin. Nuc.Acid.Res. 12:3873-3893

\bibitem{label21} Kozak,M.(1986) Influences of mRNA secondary structure
	on initiation by eucaryotic ribosomes. Proc.Natl.Acad.Sci. 83:2850-2854

\bibitem{label3} Kozak,M.(1987) An analysis of 5'-noncoding sequences from 699 vertebrate messenger RNAs. Nuc.Acid.Res. 15:8125-8148

\bibitem{label18} Kozak,M.(1987) Effects of intercistronic length on the
	efficiency of reinitiation by eucaryotic ribosomes.
	Mol.Cell.Biol. 10:3438-3445

\bibitem{label6} Kozak,M.(1989) Context effects and inefficient
initiation at non-AUG codons in eucaryotic cell-free translation
systems. Mol. Cell Biol. 9:5073-5080 

\bibitem{label4} Kozak,M.(1989) The scanning model for translation:An
	update. J.Cell.Biol. 108:229-241

\bibitem{label2} Kozak,M.(1990) Downstream secondary structure
	facilitates recognition of initiator codons by eukaryotic
	ribosomes. Proc.Natl.Acad.Sci.USA 87:8301-8305

\bibitem{label19} Kozak,M.(1991) A short leader sequence impairs the fidelity of initiation by eukaryotic ribosomes. Gene Expr. 2:111-115

\bibitem{label12} Kozak,M.(1992) Regulation of translation in eucaryotic
	systems. Annu.Rev.Cell Biol.8:197-225

\bibitem{label1} Kozak,M.(1995) Adherence to the first-AUG rule when a
	second AUG codon follows closely upon the first. Proc.Natl.Acad.Sci.USA 92:2662-2666

\bibitem{label1025} Kozak, M.(1996) Interpreting cDNA sequences: some insights from studies on translation. Mammalian Genome 7:563-574

\bibitem{label1020} Kozak, M.(1999) Initiation of translation in prokaryotes and eukaryotes. Gene 234:187-208

\bibitem{label2014} Kyrpides, N.C. and Woese, C.R.(1998) 
Archaeal translation initiation revisited: the initiation factor 2 and
	eukaryotic initiation factor 2B $\alpha$ - $\beta$ - $\delta$ subunit families. 
        Proc.Natl.Acad.Sci.USA 95:3726-3730 

\bibitem{label2015} Kyrpides N.C. and  Woese, C.R.(1998) 
  Universally conserved translation initiation factors. 
  Proc.Natl.Acad.Sci.USA 95:224-228 


\bibitem{lts14} Link, A.J., Robison, K. and Church, G.M.(1997)
	Comparing the predicted and observed properties of proteins
	encoded in the genome of {\it Escherichia coli}. Electrophoresis
	18:1259-1313

\bibitem{mglabel12} Loechel S., Inamine J.M. and Hu P-C.(1991) A novel
	translation initiation region from {\it Mycoplasma genitalium}
	that functions in {\it Escherichia coli}, Nuc.Acid.Res. 19:6905-6911

\bibitem{label31} Luukkonen, B.G., Tan, W. and Schwartz, S.(1995)
	Efficiency of reinitiation of translation on human
	immunodeficiency virus type 1 mRNAs is determined by the length
	of the upstream open reading frame and by intercistronic
	distance. J.Virol. 69:4086-4094

\bibitem{label2095} Maidak, B.L., Cole, J.R., Parker, C.T., Jr, 
Garrity, G.M., Larsen, N., Li, B., Lilburn, T.G., 
McCaughey, M.J., Olsen, G.J., Overbeek, R., Pramanik, S., 
Schmidt, T.M., Tiedje J.M. and Woese, C.R.(1999)
A new version of the RDP(Ribosomal Database Project). 
Nuc.Acid.Res. 27:171-173 

\bibitem{label7098} Maroun, L.E., Degner, M., Precup, J.W. and
Franciskovich, P.P.(1986) Eukaryotic mRNA 5'-leader sequences have
dual regions of complementarity to the 3'-terminus of 18S rRNA.
J.Theor.Biol. 120:85-98

\bibitem{label29} Mehdi, H., Ono, E. and Gupta, K.C.(1990) Initiation of translation at CUG, GUG and ACG codons in mammalian cells. Gene 91:173-178

\bibitem{label5551} Miyasaka, H.(1999) The positive relationship
between codon usage bias and translation initiation AUG context in
{\it Saccharomyces cerevisiae}. Yeast 15:633-637

\bibitem{label1939} Muggleton, S.H.(1995) Inverse entailment and
	Progol. New Gener. Comput. 13:245-286

\bibitem{label905} Nair, T.M.(1997) Calliper Randomization: An
artificial neural network based analysis of {\it E.coli} ribosome
binding sites. J.Biomol.Struct.Dynamics 15:611-617

\bibitem{arclabel_osada} Osada,Y., Saito,R. and Tomita, M.(1999) Analysis of
	base-pairing potentials between 16S rRNA and 5'UTR for
	translation initiation in various procaryotes, Bioinformatics
	15:578-581

\bibitem{label601} Pesole, G., Prunella, N., Liuni, S., Attimonelli,
M. and Saccone, C.(1992) WORDUP: an efficient algorithm for
discovering statistically significant patterns in DNA
sequences. Nuc.Acid.Res. 20:2871-2875

\bibitem{label1155} Pestova, T.V., Hellen, C.U. and Shatsky, I.N.(1996)
	Canonical eukaryotic initiation factors determine initiation of
	translation by internal ribosomal entry. Mol Cell Biol 16:6859-69

\bibitem{label1120} Pestova, T.V., Borukhov, S.I. and Hellen C.U.T.(1998) Eukaryotic ribosomes require initiation factors 1 and 1A to locate initiation codons. Nature 394:854-859

\bibitem{label10} Rinke-Appel, J., Junke, N., Brimacombe, R., Lavrik,
	I., Dokudovskaya, S., Dontsova, O. and Bogdanov, A.(1994)
	Contacts between 16S ribosomal RNA and mRNA, within the spacer
	region separating the AUG initiator codon and the Shine-Dalgarno
	sequence; a site directed cross-linking
	study. Nuc.Acid.Res. 22:3018-3025

\bibitem{label1502} Rocha, E.P.C., Danchin, A. and Viari, A.(1999)
	Translation in {\it Bacillus subtilis}: roles and trends of
	initiation and termination, insights from a genome
	analysis. Nuc.Acid.Res. 27:3567-3576

\bibitem{label659} Rudd, K.E. and Schneider, T.D.(1992) Compilation
of {\it E.coli} ribosome binding sites. in A Short course in Bacterial
Genetics : A Laboratory Manual and Handbook for {\it Escherichia coli}
and related bacteria. Cold Spring Harbor Laboratory Press, Cold Spring
Harbor, NY. 17.19-17-45

\bibitem{myco_taa} Saito, R. and Tomita, M.(1998) Analysis of translation
initiation sites in the complete genome of {\it Mycoplasma genitalium}
suggests that this bacterium uses alternative signal for translation
initiation. Genome Informatics 1998 262-263

\bibitem{neg_atg1} Saito,R. and Tomita, M.(1999) On negative selection
against ATG triplets near start codons in eukaryotic and procaryotic
genomes. J.Mol.Evol. 48:213-217.

\bibitem{neg_atg2} Saito,R. and Tomita, M.(1999)
Computer analyses of complete genomes suggest that some
archaebacteria employ both eukaryotic and eubacterial mechanisms in
translation initiation. Gene 238:79-83

\bibitem{lnis1} Salamov, A.A., Nishikawa, T. and Swindells, M.B.(1998)
	Assessing protein coding region integrity in cDNA sequencing
	projects. Bioinformatics 14:384-390

\bibitem{label7765} Salzberg, S.L.(1997) A method for identifying splice 
	sites and translational start sites in eukaryotic mRNA. CABIOS
	13:365-376

\bibitem{label4001} Sarabhai A. and Brenner, S.(1967) A mutant which
	reinitiates the polypeptide chain after chain
	termination. J.Mol.Biol. 14:145-162

\bibitem{lsaz1} Sazuka T. and Ohara, O.(1996) Sequences features surrounding the translation initiation sites assigned on the Genome Sequence of {\it Synechocystis} sp. Strain PCC6803 by amino-terminal protein sequencing. DNA Res. 3:225-232

\bibitem{label11} Schneider, T., Stormo, G.D. and Gold, L.(1986) Information content of binding sites on nucleotide sequences. J.Mol.Biol. 188:415-431

\bibitem{arclabelts200} Schneider, T.(1997) Information content of
	Individual Genetic Sequences, J.theor.Biol. 189:427-441

\bibitem{label4700} Schurr, T., Nadir, E. and Margalit, H.(1993) Identification and characterization of {\it E.coli} ribosomal binding sites by free energy computation. Nuc.Acid.Res. 21:4019-4023

\bibitem{label7} Shine, J. and Dalgarno,L.(1974) The 3'-terminal sequence of Escherichia coli 16S ribosomal RNA:Complementarity to nonsense triplets and ribosome binding sites. Proc.Nat.Acad.Sci.USA 71:1342-1346

\bibitem{label1154} Sizova D.V., Kolupaeva V.G., Pestova T.V., Shatsky
	I.N. and Hellen C.U.(1998) Specific interaction of eukaryotic
	translation initiation factor 3 with the 5' nontranslated
	regions of hepatitis C virus and classical swine fever virus
	RNAs. J.Virol. 72:4775-82 

\bibitem{label5} Slusher, L.B., Gillman, E.C., Martin, N.C. and Hopper,
	A.K.(1991) mRNA leader length and initiation codon context
	determine alternative AUG selection for the yeast gene MOD5. 
Proc.Natl.Acad.Sci.USA 88:9789-9793

\bibitem{label1205} Spanjaard, R.A. and van Duin, J.(1989) Translational reinitiation in the presence and absence of a Shine and Dalgarno sequence. Nuc.Acid.Res. 17:5501-5507

\bibitem{label1101} Sprengart, M.L., Fatscher, H.P. and Fuchs, E.(1990) The initiation of translation in {\it E.coli}: apparent base pairing between the 16srRNA and downstream sequences of the mRNA. Nuc.Acid.Res. 18:1719-1723

\bibitem{ldbox3} Sprengart, M., Fuchs, E. and Porter, A.G.(1996) The downstream
	box:An efficient and independent translation initiation signal in
	{\it Escherichia coli}. EMJO J. 15:665-674

\bibitem{arclabel5} Stormo,D., Schneider,T. and Gold,L.(1982)
	Characterization of translation initiation sites in {\it
	E.coli}. Nuc.Acid.Res. 10:2971-2996

\bibitem{label800} Stormo, G.D., Schneider, T.D., Gold, L. and
	Ehrenfeucht A.(1982) Use of 'Perceptron' algorithm to
	distinguish translational initiation sites in {\it
	E.coli}. Nuc.Acid.Res. 10:2997-3011

\bibitem{label7551} Tranque, P., Hu, M.C.-Y., Edelman, G.M. and Mauro,
V.P.(1998) rRNA complementarity within mRNAs: A possible basis for
mRNA-ribosome interactions and translational
control. Proc.Natl.Acad.Sci.USA 95:12238-12243

\bibitem{label1091} Thanaraj, T.A. and Pandit, M.W.(1989) An
additional ribosome-binding site on mRNA of highly expressed genes and
bifunctional site on the colicin fragment of 16S rRNA from {\it
Escherichia coli}: important determinants of the efficiency of
translation-initiation. Nuc.Acid.Res. 17:2973-2985

\bibitem{label532} Tomita, M., Wada, M. and Kawashima, Y.(1999) ApA
	dinucleotide periodicity in prokaryote, eukaryote, and organelle 
	genomes. J.Mol.Evol. 49:181-192.

\bibitem{label509} Tompa, M.(1999) An exact method for finding short motifs in sequences, with application to the ribosome binding site problem. ISMB'99 262-271


\bibitem{lts11} Tzareva N.V. et al.(1994) Ribosome-messenger recognition in the absence of the Shine-Dalgarno interactions. FEBS Lett. 337:189-194

\bibitem{arclabel11} Vellanoweth, R.L. and Rabinowitz, J.C.(1992) The influence
	of ribosome-binding-site elements on translational efficiency in
	{\it Bacillus subtilis} and {\it Escherichia coli} in
	vivo. Mol.Microbiol. 6:1105-1114

\bibitem{label2025} Watanabe, H., Gojobori, T. and Miura, K.(1997)
Bacterial features in the genome of {\it Methanococcus jannaschii} in
	terms of gene composition
and biased base composition in ORFs and their surrounding
regions. Gene 205:7-18

\bibitem{arclabel12} Weyens, G., Charlier, D., Roovers, M., Pierard, A. and Glansdorff, N.(1988) On the role of
the Shine-Dalgarno sequence in determining the efficiency of
translation initiation at a weak start codon in the {\it car}
operon of {\it Escherichia coli} K12. J.Mol.Biol. 204:1045-1048

\bibitem{label5050} Wilks, S.S.(1962) Testing Parametric Statistical Hypothesis. {\it In} Mathematical Statistics 394-427, John Wiley \verb+&+ Sons

\bibitem{label2099} Woese. C.R.(1998) A manifesto for microbial
genomics. Curr.Biol. 8:R781-783

\bibitem{label603} Wolfertstetter, F., Frech, K., Herrmann, G. and
Werner, T.(1996) Identification of functional elements in unaligned
nucleic acid sequences by a novel tuple search
algorithm. Bioinformatics 12:71-80

\bibitem{label515} Yada, T., Sazuka, T. and Hirosawa, M.(1997) Analysis of sequence patterns surrounding the translation initiation sites on Cyanobacterium Genome using the Hidden Markov Model. DNA Res. 4:1-7

\bibitem{label30} Yun, D.F., Laz, T.M., Clements, J.M. and Sherman,
	F.(1996) mRNA sequences influencing translation and the
	selection of AUG initiator codons in the yeast {\it
	Saccharomyces cerevisiae}. Mol.Microbiol. 19:1225-1239

\end{thebibliography}

\appendix

\chapter{Statistical backgrounds}

\input{dp_ap_stats.tex}

\chapter{Glossary}

\input{dp_ap_gloss.tex}

\end{document}




