
\begin{description}
\item[amino acid] Organic molecule containing both an amino group and a
	   carboxyl group. When more than one amino acid are linked by
	   peptide bonds, they will be protein
\item[base] Nitrogen-containing ring compounds. A,T/U,C or G.
\item[codon] Trinucleotide in a DNA or mRNA that will be translated into 
	   one specific amino acid.
\item[consensus sequence] The sequence that occurs the most frequent in specific position or for specific signal.
\item[DNA] Deoxyribonucleic acid. Polynucleotide formed from covalently linked deoxyribonucleotide units. It encodes genetic information.
\item[nucleic acid] RNA or DNA.
\item[nucleotide] = base + sugar + phosphate
\item[protein] The major macromolecular constituent of cells. A linear
	   polymer of amino acids linked together by peptide bonds in
	   a specific sequence.
\item[ribosome] Particle composed of ribosomal RNAs and ribosomal proteins that associates with mRNA and catalyzes the synthesis of protein.
\item[RNA] Ribonucleic acid. Polymer formed from covalently linked ribonucleotide monomers.
   \begin{description}
   \item[mRNA] Messenger RNA. RNA molecule that specifies the amino acid sequence of a protein. It is transcribed from DNA.
   \item[rRNA] Ribosomal RNA. Any one of a number of specific RNA molecules that form part of the structure of a ribosome and participate in the synthesis of proteins.
   \item[tRNA] Transfer RNA. Set of small RNA molecules used in protein synthesis as an interface between mRNA and amino acids.
   \end{description}
\end{description}
